---
title: "testing_algs"
author: "Fran Beresford Lynsie Warr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("mvtnorm")
library(mvtnorm)
library(invgamma)
library(truncnorm)
library(circlize)
library(ggplot2)
setwd("~/230_FINAL_PROJ")
source("algorithm_function_no_par.R")
source("optim_abc.R")
source("plot_y_trace.R")
source("plot_theta_trace.R")
```

## R Markdown

Start by simulating spatial data. 

```{r cars}
# Simulate the data

xycoords <- seq(from=0,to=1,length=10)
allcoords <- expand.grid(xycoords,xycoords)
nlocs <- nrow(allcoords)

# make the covariance function for the GPs
true.range <- 10 ## this is what they call alpha 
true.sigma2 <- 0.5
distmat <- as.matrix(dist(allcoords,diag=TRUE,upper=TRUE))
Sigma <- true.sigma2 * exp(-true.range*distmat)

# simulate the GP at each location
ys <- chol(Sigma)%*%rnorm(nlocs) # leave it at mean=0
  
# simulate poisson process based on that

#ts <- runif(nlocs,min=1,max=10) # observation durations
ts <- rep(100, nlocs)
beta <- 1 # will not be differentiated from the mean of the GP,so use mean=0
lambdas <- ts*exp(ys+beta)

Xs <- rpois(nlocs,lambdas)

# plot heat map of the data
plotdata <- data.frame(allcoords,Xs,ys)
ggplot(plotdata, aes(Var1, Var2, fill= Xs)) + 
  geom_tile()

# we want to model y, etc from above data to compare to true surface
ggplot(plotdata, aes(Var1, Var2, fill= ys)) + 
  geom_tile()
```

## testing algorithm code 

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

test <- mcmc_alg_meth_1(num_locs=nlocs, num_its= 1000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-30,beta_prop_sigma = 0.01, theta_prop_sigma = c(1,5))


#currently this does converge to values of a similar value to what we want but once we get there it doesn't move at all 
plot_y_trace(test)


## looking for 10 in left and 0.5 in right 
plot_theta_trace(test)

#should be 1
plot(test$beta, type = "l" )
```

```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-30,beta_prop_sigma = 0.01, theta_prop_sigma = c(1,5))

```



```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-50,beta_prop_sigma = 0.01, theta_prop_sigma = c(4,7))

```









```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-50,beta_prop_sigma = 0.01, theta_prop_sigma = c(1,5))

```




```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-100,beta_prop_sigma = 0.01, theta_prop_sigma = c(1,5))

```



```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-50,beta_prop_sigma = 0.0001, theta_prop_sigma = c(1,5))

```







```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 5000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-50,beta_prop_sigma = 0.5, theta_prop_sigma = c(10,15))


## looking for 10 in left and 0.5 in right in theta 
## beta is 1
```










```{r}
#mcmc_alg_meth_1(num_locs=nlocs, num_its= 10, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta = 0.01 ,beta_prop_sigma = 0.5, theta_prop_sigma = c(0.5, 0.5))

alg_and_plots(num_locs=nlocs, num_its= 10000, dist_mat = distmat, X=Xs, t_cov= ts, abc_delta= 1e-50,beta_prop_sigma = 0.5, theta_prop_sigma = c(10,15))
## looking for 10 in left and 0.5 in right in theta 
## beta is 1
```



It seems that as soon as y gets close enough to the true values, it stops mixing, sometimes this is after it has moved 30 times but sometimes this occurs after only one move. 

Beta has been the easiest to get closest to its true value. It seems that 0.01 is a good value for the proposal sigma regardless of what the other parameters are doing.

The theta values are not being found at all. 